Analysis of competitive surfing tournaments with generalized Bradley-Terry likelihoods

### 2022-04-26 inst/surfing_jsa_rebuttal_referee2.tex

\section*{MS596 ``Analysis of competitive surfing tournaments
  with generalized Bradley-Terry likelihoods" by Driver and
  Hankin: second rebuttal}

Below, the reviewer's comments are in black, and our replies to the
issues are in \textcolor{blue}{blue}.  We have indicated changes to
the manuscript where appropriate.

Short story: we have accommodated all the comments with rewording and
clarification.

Slightly longer story: the referee makes a wide range of observations,
all which we have engaged with.  We have overhauled much of the
writing, added descriptions of the computational and mathematical
methodology, clarified ambiguous wording, produced new statistical
tests, and improved the figures.  We have tried hard to retain the
many positive aspects of the manuscript that the referee explicitly
notes.

@@

We have included an extended discussion of some of the points below.
The resulting document is, we believe, stronger and more scholarly
than before and we recommend it to you.

\section*{Referee's comments}

The paper does a good job of outlining the basics of surfing in terms
of competition rules and structure, and frames up the problem at
hand.  However, the paper itself feels rather light on content.  The
authors jump almost immediately to the results section and don't go
into very much detail in terms of how or why Bradley-Terry models are
used, other than the fact that they are used.

\textcolor{blue}{We have re-read our manuscript and indeed
  Bradley-Terry (BT) does make a rather abrupt entrance.  However, we
  do describe and justify and motivate the use of BT in section three.
  On reflection we believe we can address this comment by rearranging
  sections 2 and 3: this gives a gentler and more appropriate
  introduction to BT and indeed is a more appropriate structure for a
  manuscript.  In line with this, we have changed the section heading
  from ``previous related work'' to ``Previous statistical analysis of
  competitive surfing'' which is a better description, we feel.  We
  have added several new statistical tests and given brief outlines of
  their significance.}

@@

From a style perspective, the paper is well-written and enjoyable to
read, which makes the lack of substantial content a bit of a shame, as
the authors are clearly able to write a paper worth reading. Overall,
it sounds like the authors do have the data and analysis available to
warrant a more complete paper, however, what they have submitted feels
rushed and rather rough around the edges. While possible, it would
take a fairly significant overhaul/expansion to the current version of
the paper to get it into a publication-ready format.

\textcolor{blue}{We believe that this referee's insightful and
  constructive comments have led to a much improved submission.}

### 2022-04-26 inst/surfing_jsa.tex
tweaks

The 2019 WSL tour saw competitive events held at eleven
different surf locations.

The typical format is as follows.  Up to four surfers---the
competitors---are in the water simultaneously, watched by up to five
judges.  Subject to surfing etiquette, competitors are free to catch a
wave at their discretion within the heat's time window.  The judges
rate each wave ride and award points to the surfer based on a range of
subjective criteria such as stylish execution and transition of
manoeuvres, but also credit elements such as the difficulty and
novelty of the performance.  The top score and bottom score among the
five judges are removed and the remaining three judges' scores are
averaged to give the final score for the surfer for that wave.  Each
surfer's aggregate score is given by the average of the two
highest-scoring waves.  The winner is the surfer with the highest
aggregate score.

@@

\begin{equation}
  \frac{\partial^2}{\partial p_i\partial p_j}
  \log\left(\sum p_k\right)^n=
  \begin{cases}
    -n/\left(p_i+p_j+\sum p_k\right)^2\qquad\mbox{if $i,j\in K$}\\
    0\qquad\mbox{otherwise.}
    \end{cases}
\end{equation}

@@

Consider, as an example, the case of Smith, Buchan, Freestone who were
simultaneously in the water at Bells Beach, Australia.  Now, Buchan
and Freestone are Australian, and Jordy Smith is South African.  It is
reasonable to ask whether Buchan and Freestone, who are in direct
competition with one another for a place in the Australian Olympic
surfing team, would be more motivated to beating each other than to
come ahead of Smith.  This would be a rational response, but arguably
noncompetitive.  How would such a strategy manifest itself in
scorelines?  We would suggest that Smith is able to modify his
strategy to benefit from the mutual rivalry between Buchan and
Freestone; and indeed we see Smith winning that particular heat.  We
tentatively interpret this as a Nash-type equilibrium: there is no
incentive for either Buchan nor Freestone to deviate from their
prioritization of defeating their compatriot; and indeed neither
Buchan nor Freestone were selected for the 2020 Olympic team.

>>>>

The 2019 WSL tour saw competitive events held at eleven different surf
locations.  The typical format is as follows.  Up to four
surfers---the competitors---are in the water simultaneously, watched
by up to five judges.  Subject to surfing etiquette, competitors are
free to catch a wave at their discretion within the heat's time
window.  The judges rate each wave ride and award points to the surfer
based on a range of subjective criteria such as stylish execution and
transition of manoeuvres, but also credit elements such as the
difficulty and novelty of the performance.  The top score and bottom
score among the five judges are removed and the remaining three
judges' scores are averaged to give the final score for the surfer for
that wave.  Each surfer's aggregate score is given by the average of
the two highest-scoring waves.  The winner is the surfer with the
highest aggregate score.

@@

\begin{equation}
  \frac{\partial^2}{\partial p_i\partial p_j}
  \log\left(\sum p_k\right)^n=
  \begin{cases}
    -n/\left(\sum p_k\right)^2\qquad\mbox{if $i,j\in K$}\\
    0\qquad\mbox{otherwise.}
    \end{cases}
\end{equation}

@@

Consider, as an example, the case of Smith, Buchan, Freestone who were
simultaneously in the water at Bells Beach, Australia.  Now, Buchan
and Freestone are Australian, and Jordy Smith is South African.  It is
reasonable to ask whether Buchan and Freestone, who are in direct
competition with one another for a place in the Australian Olympic
surfing team, would be more motivated to beat each other than to
come ahead of Smith.  This would be a rational response, but arguably
noncompetitive.  How would such a strategy manifest itself in
scorelines?  We would suggest that Smith is able to modify his
strategy to benefit from the mutual rivalry between Buchan and
Freestone; and indeed we see Smith winning that particular heat.  We
tentatively interpret this as a Nash-type equilibrium: there is no
incentive for either Buchan nor Freestone to deviate from their
prioritization of defeating their compatriot; and indeed neither
Buchan nor Freestone were selected for the 2020 Olympic team.
